---
title: "Astrocyte (GFAP) Analysis"
author: "Vatsal D. Jariwala"
date: "2024-10-24"
output: html_document
---

## Introduction

This section visualizes GFAP-labeled astrocyte data across AAV-transfected samples. The analyses include:
- Astrocyte % relative to DAPI
- Total astrocyte counts
- Colocalization with AAV expression (ROI)
- Total DAPI counts for normalization context

---

## 1. Load Required Packages

```{r message=FALSE}
library(tidyverse)

if (!requireNamespace("Seurat", quietly = TRUE)) {
  install.packages("Seurat")
}
library(Seurat)
```

---

## 2. Import GFAP Data

```{r}
df.GFAP <- na.omit(
  readxl::read_excel(
    "path to respective cell count file",
    sheet = "df_GFAP_avg"
  )
)
```

---

## 3. Astrocyte Percentage (% DAPI)

```{r}
summary_stats <- df.GFAP %>%
  group_by(Timepoint, Transfection) %>%
  summarise(
    mean_percent = mean(Avg_strocyte_Percent_DAPI, na.rm = TRUE),
    se_percent = sd(Avg_strocyte_Percent_DAPI, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(summary_stats, aes(x = factor(Timepoint, levels = c("DIV5", "DIV8")),
                          y = mean_percent,
                          fill = Timepoint)) +
  geom_bar(stat = "identity", position = position_dodge(0.9), color = "black") +
  geom_errorbar(aes(ymin = mean_percent - se_percent, ymax = mean_percent + se_percent),
                width = 0.2, position = position_dodge(0.9)) +
  facet_grid(~ factor(Transfection, levels = c("Control_PBS", "CMV", "Hsyn")), scales = "free_y") +
  labs(y = "Astrocyte Percentage (% DAPI)", x = "") +
  ylim(0, 100) +
  theme_bw() +
  NoLegend()
```

---

## 4. Absolute Astrocyte Count

```{r}
summary_stats <- df.GFAP %>%
  group_by(Timepoint, Transfection) %>%
  summarise(
    mean_percent = mean(Avg_Astrocyte, na.rm = TRUE),
    se_percent = sd(Avg_Astrocyte, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(summary_stats, aes(x = factor(Timepoint, levels = c("DIV5", "DIV8")),
                          y = mean_percent,
                          fill = Timepoint)) +
  geom_bar(stat = "identity", position = position_dodge(0.9), color = "black") +
  geom_errorbar(aes(ymin = mean_percent - se_percent, ymax = mean_percent + se_percent),
                width = 0.2, position = position_dodge(0.9)) +
  facet_grid(~ factor(Transfection, levels = c("Control_PBS", "CMV", "Hsyn")), scales = "free_y") +
  labs(y = "Astrocyte Count", x = "") +
  ylim(0, 100) +
  theme_bw() +
  NoLegend()
```

---

## 5. Astrocyte Colocalization with AAV (ROI)

```{r}
summary_stats <- df.GFAP %>%
  group_by(Timepoint, Transfection) %>%
  summarise(
    mean_percent = mean(Avg_GFAP_AAV, na.rm = TRUE),
    se_percent = sd(Avg_GFAP_AAV, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(summary_stats, aes(x = factor(Timepoint, levels = c("DIV5", "DIV8")),
                          y = mean_percent,
                          fill = Timepoint)) +
  geom_bar(stat = "identity", position = position_dodge(0.9), color = "black") +
  geom_errorbar(aes(ymin = mean_percent - se_percent, ymax = mean_percent + se_percent),
                width = 0.2, position = position_dodge(0.9)) +
  facet_grid(~ factor(Transfection, levels = c("Control_PBS", "CMV", "Hsyn")), scales = "free_y") +
  labs(y = "Astrocyte Colocalization (ROI)", x = "") +
  ylim(0, 100) +
  theme_bw() +
  NoLegend()
```

---

## 6. Total DAPI Counts

```{r}
summary_stats <- df.GFAP %>%
  group_by(Timepoint, Transfection) %>%
  summarise(
    mean_percent = mean(Avg_DAPI, na.rm = TRUE),
    se_percent = sd(Avg_DAPI, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(summary_stats, aes(x = factor(Timepoint, levels = c("DIV5", "DIV8")),
                          y = mean_percent,
                          fill = Timepoint)) +
  geom_bar(stat = "identity", position = position_dodge(0.9), color = "black") +
  geom_errorbar(aes(ymin = mean_percent - se_percent, ymax = mean_percent + se_percent),
                width = 0.2, position = position_dodge(0.9)) +
  facet_grid(~ factor(Transfection, levels = c("Control_PBS", "CMV", "Hsyn")), scales = "free_y") +
  labs(y = "DAPI Count", x = "") +
  theme_bw() +
  NoLegend()
```

---

## Notes

- File paths are local; ensure correct directory if moving between systems.
- Output is consistent with the Gfap analysis for clear comparison.
- This document is intended for collaborative visualization and version control on GitHub.
