---
title: "LDH Assay Data Visualization"
author: "Vatsal D. Jariwala"
date: "2024-10-24"
output: html_document
---

## Introduction

This R Markdown document analyzes and visualizes LDH (Lactate Dehydrogenase) levels under different experimental conditions in an AAV project. It includes data import from Excel, package installation, and data visualization using `ggplot2`.

---

## 1. Package Installation and Loading

```{r message=FALSE}
# Create a list object to store package information
packages <- list()

# Specify all required packages
packages$my_packages <- c("readr", "plyr", "readxl", "dplyr", 
                          "factoextra", "tidyverse", "tidyr", 
                          "ggplot2", "stringr")

# Identify any packages that are not installed
packages$not_installed <- packages$my_packages[!(packages$my_packages %in% installed.packages()[, "Package"])]

# Install only missing packages
if(length(packages$not_installed)) {
  install.packages(packages$not_installed)
}

# Load all required packages
lapply(packages$my_packages, require, character.only = TRUE)
```

---

## 2. Data Import

```{r}
# Create a list to store data frames
import <- list()

# Load the LDH assay dataset from the specified Excel file
import$df_LDH <- read_excel("/Users/vatsaljariwala/Documents/AAV_Project/LDH assay/Datasheet/LDH_data_all.xlsx", 
                            sheet = "Sheet1")
```

---

## 3. LDH Levels by Condition (All Patients)

```{r}
# Plot showing LDH levels over time for each condition
ggplot(import$df_LDH, aes(x = (DIV), y = LDH_level, fill = Condition, color = Condition)) +
  geom_point(aes(shape = factor(Patient))) +                      # Plot individual data points
  geom_smooth(se = FALSE, method = "loess", linewidth = 0.5) +   # Add LOESS trend line
  theme_classic() +
  xlab("DIV") +
  ylab("LDH Levels (U/ml)") +
  labs(color = "Condition", fill = "Condition", shape = "Patient ID") +
  scale_fill_manual(values = c("red", "#00AFBB", "green")) +
  scale_color_manual(values = c("red", "#00AFBB", "green")) +
  scale_x_discrete(limits = as.character(0:9)) +                 # Ensure consistent x-axis labels
  theme(
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 9, face = "bold"),
    legend.title = element_text(size = 9, face = "bold"),
    legend.key.size = unit(1, "lines"),
    legend.position = "right",
    strip.text = element_text(size = 10, face = "bold")
  )
```

---

## 4. LDH Levels by Patient and Condition (Faceted)

```{r}
# Same plot, now faceted by Patient
ggplot(import$df_LDH, aes(x = (DIV), y = LDH_level, fill = Condition, color = Condition)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "loess") +
  theme_classic() +
  xlab("DIV") +
  ylab("LDH Levels (U/ml)") +
  labs(color = "Condition", fill = "Condition", shape = "Patient ID") +
  scale_fill_manual(values = c("red", "#00AFBB", "green")) +
  scale_color_manual(values = c("red", "#00AFBB", "green")) +
  facet_grid(~Patient) +                                          # Create a plot per patient
  scale_x_discrete(limits = as.character(0:9)) +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 9, face = "bold"),
    legend.title = element_text(size = 9, face = "bold"),
    legend.key.size = unit(1, "lines"),
    legend.position = "right",
    strip.text = element_text(size = 10, face = "bold")
  )
```

---

## Notes

- Ensure that the Excel file path is correctly specified on your system.
- If the `DIV` variable is numeric and you want discrete behavior, convert it to character or factor using `as.character()` or `as.factor()`.
- This analysis can be extended with statistical comparisons, violin plots, and outlier detection depending on further research needs.
