---
title: "NeuN Neuron Analysis"
author: "Kevin Joseph, Vatsal D. Jariwala"
date: "2024-10-24"
output: html_document
---

## Introduction

This R Markdown file visualizes neuronal counts labeled by NeuN in AAV-transfected cultures. Similar to the Iba1 analysis, we assess:
- Neuron percentage relative to DAPI
- Absolute NeuN+ cell count
- NeuN colocalization in AAV-expressing ROIs

All data is grouped by `Timepoint` and `Transfection` for consistency and comparability.

---

## 1. Load Required Packages

```{r message=FALSE}
# Load essential package
library(tidyverse)

# For removing legends from ggplot
if (!requireNamespace("Seurat", quietly = TRUE)) {
  install.packages("Seurat")
}
library(Seurat)
```

---

## 2. Import Data

```{r}
# Read the Excel data and omit missing values
df.Neun <- na.omit(
  readxl::read_excel(
    "path to respective cell count file",
    sheet = "df_Neun_avg"
  )
)
```

---

## 3. Plot Neuron Percentage (% DAPI)

```{r}
# Calculate summary statistics
summary_stats <- df.Neun %>%
  group_by(Timepoint, Transfection) %>%
  summarise(
    mean_percent = mean(Avg_Neuron_Percent_DAPI, na.rm = TRUE),
    se_percent = sd(Avg_Neuron_Percent_DAPI, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Generate bar plot
ggplot(summary_stats, aes(x = factor(Timepoint, levels = c("DIV5", "DIV8")),
                          y = mean_percent,
                          fill = Timepoint)) +
  geom_bar(stat = "identity", position = position_dodge(0.9), color = "black") +
  geom_errorbar(aes(ymin = mean_percent - se_percent, ymax = mean_percent + se_percent),
                width = 0.2, position = position_dodge(0.9)) +
  facet_grid(~ factor(Transfection, levels = c("Control_PBS", "CMV", "Hsyn")), scales = "free_y") +
  labs(y = "Neuron Percentage (% DAPI)", x = "") +
  ylim(0, 100) +
  theme_bw() +
  NoLegend()
```

---

## 4. Plot Absolute Neuron Counts

```{r}
# Calculate summary statistics
summary_stats <- df.Neun %>%
  group_by(Timepoint, Transfection) %>%
  summarise(
    mean_percent = mean(Avg_Neun, na.rm = TRUE),
    se_percent = sd(Avg_Neun, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Generate bar plot
ggplot(summary_stats, aes(x = factor(Timepoint, levels = c("DIV5", "DIV8")),
                          y = mean_percent,
                          fill = Timepoint)) +
  geom_bar(stat = "identity", position = position_dodge(0.9), color = "black") +
  geom_errorbar(aes(ymin = mean_percent - se_percent, ymax = mean_percent + se_percent),
                width = 0.2, position = position_dodge(0.9)) +
  facet_grid(~ factor(Transfection, levels = c("Control_PBS", "CMV", "Hsyn")), scales = "free_y") +
  labs(y = "Neuron Count", x = "") +
  ylim(0, 100) +
  theme_bw() +
  NoLegend()
```

---

## 5. Plot Neuron Colocalization (ROI)

```{r}
# Calculate summary statistics
summary_stats <- df.Neun %>%
  group_by(Timepoint, Transfection) %>%
  summarise(
    mean_percent = mean(Avg_Neun_AAV, na.rm = TRUE),
    se_percent = sd(Avg_Neun_AAV, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Generate bar plot
ggplot(summary_stats, aes(x = factor(Timepoint, levels = c("DIV5", "DIV8")),
                          y = mean_percent,
                          fill = Timepoint)) +
  geom_bar(stat = "identity", position = position_dodge(0.9), color = "black") +
  geom_errorbar(aes(ymin = mean_percent - se_percent, ymax = mean_percent + se_percent),
                width = 0.2, position = position_dodge(0.9)) +
  facet_grid(~ factor(Transfection, levels = c("Control_PBS", "CMV", "Hsyn")), scales = "free_y") +
  labs(y = "Neuronal Colocalization (ROI)", x = "") +
  ylim(0, 100) +
  theme_bw() +
  NoLegend()
```

---

## Notes

- File paths are local; ensure correct directory if moving between systems.
- Output is consistent with the Neun analysis for clear comparison.
- This document is intended for collaborative visualization and version control on GitHub.
