---
title: "Oligodendrocyte (Olig) Analysis"
author: "Kevin Joseph, Vatsal D. Jariwala"
date: "2024-10-24"
output: html_document
---

## Introduction

This R Markdown file visualizes oligodendrocyte counts labeled by Olig marker across AAV-transfected cultures. The analysis includes:
- Olig percentage relative to DAPI
- Absolute Olig+ cell count
- Olig colocalization in AAV-expressing ROIs

---

## 1. Load Required Packages

```{r message=FALSE}
library(tidyverse)

# For removing legends from ggplot
if (!requireNamespace("Seurat", quietly = TRUE)) {
  install.packages("Seurat")
}
library(Seurat)
```

---

## 2. Import Data

```{r}
# Read the Excel data and omit missing values
df.Olig <- na.omit(
  readxl::read_excel(
    "path to respective cell count file",
    sheet = "df_Olig_avg"
  )
)
```

---

## 3. Plot Oligodendrocyte Percentage (% DAPI)

```{r}
# Calculate summary statistics
summary_stats <- df.Olig %>%
  group_by(Timepoint, Transfection) %>%
  summarise(
    mean_percent = mean(Avg_Olig_Percent_DAPI, na.rm = TRUE),
    se_percent = sd(Avg_Olig_Percent_DAPI, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Generate bar plot
ggplot(summary_stats, aes(x = factor(Timepoint, levels = c("DIV5", "DIV8")),
                          y = mean_percent,
                          fill = Timepoint)) +
  geom_bar(stat = "identity", position = position_dodge(0.9), color = "black") +
  geom_errorbar(aes(ymin = mean_percent - se_percent, ymax = mean_percent + se_percent),
                width = 0.2, position = position_dodge(0.9)) +
  facet_grid(~ factor(Transfection, levels = c("Control_PBS", "CMV", "Hsyn")), scales = "free_y") +
  labs(y = "Oligodendrocyte Percentage (% DAPI)", x = "") +
  ylim(0, 100) +
  theme_bw() +
  NoLegend()
```

---

## 4. Plot Absolute Oligodendrocyte Counts

```{r}
# Calculate summary statistics
summary_stats <- df.Olig %>%
  group_by(Timepoint, Transfection) %>%
  summarise(
    mean_percent = mean(Avg_Olig, na.rm = TRUE),
    se_percent = sd(Avg_Olig, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Generate bar plot
ggplot(summary_stats, aes(x = factor(Timepoint, levels = c("DIV5", "DIV8")),
                          y = mean_percent,
                          fill = Timepoint)) +
  geom_bar(stat = "identity", position = position_dodge(0.9), color = "black") +
  geom_errorbar(aes(ymin = mean_percent - se_percent, ymax = mean_percent + se_percent),
                width = 0.2, position = position_dodge(0.9)) +
  facet_grid(~ factor(Transfection, levels = c("Control_PBS", "CMV", "Hsyn")), scales = "free_y") +
  labs(y = "Oligodendrocyte Count", x = "") +
  ylim(0, 100) +
  theme_bw() +
  NoLegend()
```

---

## 5. Plot Olig Colocalization (ROI)

```{r}
# Calculate summary statistics
summary_stats <- df.Olig %>%
  group_by(Timepoint, Transfection) %>%
  summarise(
    mean_percent = mean(Avg_Olig_AAV, na.rm = TRUE),
    se_percent = sd(Avg_Olig_AAV, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Generate bar plot
ggplot(summary_stats, aes(x = factor(Timepoint, levels = c("DIV5", "DIV8")),
                          y = mean_percent,
                          fill = Timepoint)) +
  geom_bar(stat = "identity", position = position_dodge(0.9), color = "black") +
  geom_errorbar(aes(ymin = mean_percent - se_percent, ymax = mean_percent + se_percent),
                width = 0.2, position = position_dodge(0.9)) +
  facet_grid(~ factor(Transfection, levels = c("Control_PBS", "CMV", "Hsyn")), scales = "free_y") +
  labs(y = "Oligodendrocyte Colocalization (ROI)", x = "") +
  ylim(0, 100) +
  theme_bw() +
  NoLegend()
```

---
## Notes

- File paths are local; ensure correct directory if moving between systems.
- Output is consistent with the Oligs analysis for clear comparison.
- This document is intended for collaborative visualization and version control on GitHub.

